### 2016-05-03 paper.tex
features of fk tables

%% This template can be used to write a paper for
%% Computer Physics Communications using LaTeX.
%% For authors who want to write a computer program description,
%% an example Program Summary is included that only has to be
%% completed and which will give the correct layout in the
%% preprint and the journal.
%% The `elsarticle' style is used and more information on this style
%% can be found at 
%% http://www.elsevier.com/wps/find/authorsview.authors/elsarticle.
%%
%%
\documentclass[preprint,12pt]{elsarticle}

\textwidth=16.0cm \textheight=21.0cm 
\topmargin 0cm \oddsidemargin 0cm 
\setlength{\unitlength}{1mm}

%% Use the option review to obtain double line spacing
%% \documentclass[preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

%% if you use PostScript figures in your article
%% use the graphics package for simple commands
%% \usepackage{graphics}
%% or use the graphicx package for more complicated commands
%% \usepackage{graphicx}
%% or use the epsfig package if you prefer to use the old commands
%% \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
\usepackage{url,hyperref}

%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%
%% \biboptions{comma,round}

% \biboptions{}
\biboptions{numbers,sort&compress}

\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\ba}{\begin{eqnarray}}
\newcommand{\ea}{\end{eqnarray}}

%% This list environment is used for the references in the
%% Program Summary
%%
\newcounter{bla}
\newenvironment{refnummer}{%
\list{[\arabic{bla}]}%
{\usecounter{bla}%
 \setlength{\itemindent}{0pt}%
 \setlength{\topsep}{0pt}%
 \setlength{\itemsep}{0pt}%
 \setlength{\labelsep}{2pt}%
 \setlength{\listparindent}{0pt}%
 \settowidth{\labelwidth}{[9]}%
 \setlength{\leftmargin}{\labelwidth}%
 \addtolength{\leftmargin}{\labelsep}%
 \setlength{\rightmargin}{0pt}}}
 {\endlist}

 \journal{Computer Physics Communications}

 \begin{document}

 \begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{{\tt APFELgrid}: a fast interface to hadronic observables for
  fits of parton densities}
\tnotetext[mytitlenote]{Preprints: OUTP-16-09P, CERN-TH-2016-103}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

\author[a]{Valerio Bertone}
\author[b]{Stefano Carrazza}
\author[a]{Nathan P. Hartland\corref{author}}

\cortext[author] {Corresponding author.\\\textit{E-mail address:} nathan.hartland@physics.ox.ac.uk}
\address[a]{Rudolf Peierls Centre for Theoretical Physics,\\ 1 Keble Road, University of Oxford, OX1 3NP, Oxford, UK}
\address[b]{Theoretical Physics Department, CERN, Geneva, Switzerland}

\begin{abstract}
  We present a new software package designed to aid the inclusion of
  hadronic observables into Parton Distribution Function (PDF)
  fits. The {\tt APFELgrid} package converts interpolated weight
  tables provided by {\tt APPLgrid} files into a more efficient format
  for PDF fitting by the pre-computation and combination with PDF and
  $\alpha_s$ evolution factors provided by {\tt APFEL}. This
  combination significantly reduces the number of operations required
  to perform the calculation of hadronic observables in PDF fits and
  simplifies the structure of the calculation into a readily optimised
  scalar product. We demonstrate that our technique can lead to a
  substantial speed improvement as compared to the existing interfaces
  without any reduction in numerical accuracy.
  % We present a new method conceived to ease the inclusion of
  % hadronic observables into parton distribution function (PDF)
  % fits. Our technique relies on the the same principles adopted by
  % the existing fast interfaces but implements a set of improvements
  % aimed at optimizing the performance in the context of PDF fits
  % where many iterations and many predictions for each iteration are
  % usually required. The main novelties are the precomputation of the
  % PDF and $\alpha_s$ evolution and the optimisation of the numerical
  % convolution between hard cross sections and PDFs. We demonstrate
  % that our technique can lead to a substantial speed improvement as
  % compared to the existing interfaces.
\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
QCD; parton distribution functions; fast predictions.
\end{keyword}

\end{frontmatter}

 \newpage

\noindent {\Large \textbf{Program Summary}}\\

\begin{small}
\noindent
{\em Manuscript Title:} {\tt APFELgrid}: a fast interface to hadronic observables for
  fits of parton densities                                      \\
{\em Authors:} V.~Bertone, S.~Carrazza, N.P.~Hartland                                               \\
{\em Program Title:} {\tt APFELgrid}                                          \\
{\em Journal Reference:}                                      \\
  %Leave blank, supplied by Elsevier.
{\em Catalogue identifier:}                                   \\
  %Leave blank, supplied by Elsevier.
{\em Licensing provisions:}   MIT license                                \\
{\em Programming language:}  C++                                 \\
{\em Computer:} PC/Mac                                               \\
  %Computer(s) for which program has been designed.
{\em Operating system:} MacOS/Linux                                       \\
  %Operating system(s) for which program has been designed.
{\em RAM:} varying                                              \\
  %RAM in bytes required to execute program with typical data.
{\em Keywords:} QCD, PDF\\
  % Please give some freely chosen keywords that we can use in a
  % cumulative keyword index.
{\em Classification:}  11.1 General, High Energy Physics and Computing                                       \\
  %Classify using CPC Program Library Subject Index, see (
  % http://cpc.cs.qub.ac.uk/subjectIndex/SUBJECT_index.html)
  %e.g. 4.4 Feynman diagrams, 5 Computer Algebra.
{\em External routines/libraries:}  {\tt APPLgrid}, {\tt APFEL}                          \\
  % Fill in if necessary, otherwise leave out.
{\em Nature of problem:}\\
 Fast computation of hadronic observables under the variation of parton distribution functions.
   \\
{\em Solution method:}\\
  Combination of interpolated weight grids from {\tt APPLgrid} and evolution factors from {\tt APFEL} into efficient {\tt FastKernel} tables.\\
{\em Running time:} varying\\
   \\
\end{small}

\clearpage

%\tableofcontents

%% main text
\section{Introduction}\label{sec:intro}

The ability to perform fast and accurate cross-section predictions for
measurements at the Large Hadron Collider (LHC) is a vital requirement
for the most precise PDF determinations. LHC measurements have a
unique capacity to shed light upon the dynamics of the proton and
provide stringent constraints upon the proton PDF. In order to make
full use of current and future LHC data, efficient strategies for the
computation of predictions for these observables must be
deployed. Software packages performing predictions for non-inclusive
hadron-collider observables cannot be easily deployed in a PDF fit
requiring rapid iteration due the time required to obtain accurate
results (usually of the order of a few hours or more per data
point). In order to overcome this limitation, the typical strategy
adopted for fast cross-section prediction relies on the precomputation
of the partonic hard cross sections in such a way that the numerical
convolution with any set of PDFs can be quickly performed by means of
interpolation techniques.

This interpolation strategy is implemented in the {\tt
  APPLgrid}~\cite{Carli:2010rw} and {\tt
  FastNLO}~\cite{Wobisch:2011ij} projects. For the computation of the
hard cross-sections, these packages rely on external codes to which
they are interfaced by means of a suite of functions allowing for the
filling of PDF- and $\alpha_s$-independent look-up tables of
cross-section weights. Codes such as {\tt MCFM}~\cite{Campbell:2010ff}
and {\tt NLOJet++}~\cite{Nagy:2003tz} have been interfaced directly to
{\tt APPLgrid}/{\tt FastNLO} and more recently dedicated interfaces to
automated general-purpose Monte Carlo (MC) event generators have been
developed. The {\tt aMCfast}~\cite{Bertone:2014zva} and {\tt
  MCgrid}~\cite{DelDebbio:2013kxa} codes have been specifically
designed to extract the relevant information from the {\tt
  MadGraph5\_aMC@NLO}~\cite{Alwall:2014hca} and {\tt
  SHERPA}~\cite{Gleisberg:2008ta} event generators respectively.

While these tools have proven to be extremely useful in the extraction
of parton densities, the volume of experimental data made available by
LHC collaborations for use in PDF fits is already stretching the
capabilities of typical fitting technology. Presently a typical global
PDF fit includes thousands of hadronic data points for which
predictions have to be computed thousands of times during the
minimisation process. As a consequence the ``standard'' tools, $i.e.$
{\tt APPLgrid} and {\tt FastNLO}, might not be fast enough to meet the
needs of modern PDF fitters. For this reason a high-performance tool
tailored specifically for PDF analyses is becoming increasingly
important.

The {\tt FastKernel} method, developed in the framework of the NNPDF
global analyses~\cite{Ball:2010de,Ball:2014uwa}, was developed to
address this problem. This method differs from the standard procedure
\`{a} la {\tt APPLgrid} or {\tt FastNLO} in that it maximises the
amount of information that is precomputed prior to fitting such as to
minimise the amount of operations to be performed during the
fit. Specifically, the {\tt FastKernel} method relies on the
precomputation of the partonic hard cross-sections and DGLAP evolution
kernels and their combination into a single look-up table, here called
a {\tt FastKernel} ({\tt FK}) table, in such a way that the prediction
for a given hadronic observable can be obtained by a simple matrix
product between the respective {\tt FK} table and the initial scale
PDFs.

In this paper we present the {\tt APFELgrid} package, a public
implementation of the {\tt FastKernel} method in which the hard
partonic cross sections provided in an {\tt APPLgrid} look-up table
are combined with the DGLAP evolution kernels provided by the {\tt
  APFEL}
package~\cite{Bertone:2013vaa,Carrazza:2014gfa,Bertone:2015cwa,Bertone:2015lqa,Carrazza:2015dea,Bertone:2015gba}.

The outline of this paper is the following. In
Sect.~\ref{sec:FastKernel} we present the technical details of the
implementation of the {\tt FastKernel} method. Then, in
Sect.~\ref{sec:benchmark} we illustrate timings and performance
results of the {\tt APFELgrid} library and grids. Finally, in
Sect.~\ref{sec:conclusion} we conclude and discuss possible future
developments.

\section{Interpolation tools for collider observables}\label{sec:FastKernel}

Hadron collider observables are typically computed in QCD by a double
convolution of parton densities with a hard scattering
cross-section. Consider for example the calculation of a general
cross-section $pp\to X$ with a set of PDFs $\{f\}$:
\begin{equation}\label{eq:hadconv}
  \sigma_{pp\to X} =
  \sum_{s}\sum_{p} \int dx_1\,dx_2\,
  \hat{\sigma}^{(p)(s)}\,\alpha_s^{p+p_{\rm LO}}(Q^2) \,F^{(s)}(x_1,x_2, Q^2)\,.
\end{equation}
where $Q^2$ is the typical hard scale of the process, the index
$s$ sums over the active partonic subprocesses in the calculation, $p_{\rm LO}$ is the leading-order power of $\alpha_s$ for the process
and $\hat{\sigma}^{(p)(s)}$ is the N$^p$LO cross-section for the partonic subprocess scattering $s\to X$. $F^{(s)}$ represents the 
subprocess parton density:
\begin{equation}\label{eq:APPLsubproc}
  F^{(s)}(x_1,x_2, Q^2) =\sum_{i,j} C^{(s)}_{ij} 
  f_i(x_{1},Q^2)f_j(x_{2},Q^2)\,,
\end{equation}
where the $C^{(s)}_{ij}$ matrix enumerates the combinations of PDFs
that contribute to the $s$-th subprocess.  The central observation of
tools such as {\tt APPLgrid} and {\tt FastNLO} is that the PDF and
$\alpha_S$ dependence may be factorised out of the convolution via
expansion over a set of interpolating functions, spanning $Q^2$ and
the two values of parton-$x$. For example one may represent the
subprocess PDFs and $\alpha_S$ in terms of the Lagrange basis
polynomials $\mathcal{I}_\tau(Q^2)$, $\mathcal{I}_\alpha(x_1)$ and
$\mathcal{I}_\beta(x_2)$ as:
\begin{equation}\label{eq:interpolation}
\begin{array}{c}
\displaystyle \alpha_s^{p+p_{\rm LO}}(Q^2) \,F^{(s)}(x_1,x_2, Q^2)
  =\\
\\
\displaystyle \sum_{\alpha,\beta,\tau} \alpha_s^{p+p_{\rm LO}}(Q_\tau^2)
  \,F^{(s)}_{\alpha\beta , \tau}
  \,\mathcal{I}_\tau(Q^2)\,\mathcal{I}_\alpha(x_1)
  \,\mathcal{I}_\beta(x_2)
\end{array}
\end{equation}
where we use the shorthand
$F^{(s)}_{\alpha\beta ,\tau} = F^{(s)}(x_\alpha, x_\beta,Q_\tau^2)$.
Using these expressions in the double-convolution of
Eq.~(\ref{eq:hadconv}) one can finally obtain an expression for the
desired cross-section which depends upon the subprocess PDFs only
though a simple product:

\begin{equation} \label{eq:applconv}
  \sigma_{pp\to X} = \sum_p^{N_\alpha} \sum_{s}^{N_{\mathrm{sub}}} \sum_{\alpha,\beta,\tau} 
  \alpha_s^{p+p_{\rm LO}}(Q^2_\tau)W_{\alpha\beta,\tau}^{(p)(s)} \, F_{\alpha\beta,\tau}^{(s)}\,.
\end{equation}
where
\begin{equation}
  W_{\alpha\beta,\tau}^{(p)(s)} = \mathcal{I}_\tau(Q^2)\int dx_1\,dx_2\,
  \hat{\sigma}^{(p)(s)}\,\mathcal{I}_\alpha(x_1)
  \,\mathcal{I}_\beta(x_2)\,,
\end{equation}
consists of the convolution of the hard cross-section with the
interpolating polynomials. This information may be stored in a
precomputed look-up table.  The final expression for the cross-section
in Eq.~(\ref{eq:applconv}) therefore being considerably more
computationally efficient inside a fit than the direct performance of
the double-convolution.

\subsection{The FastKernel method}

A number of tools, $e.g.$ {\tt
  APFEL}~\cite{Bertone:2013vaa,Carrazza:2014gfa,Bertone:2015cwa,Bertone:2015lqa,Carrazza:2015dea,Bertone:2015gba}
or {\tt HOPPET}~\cite{Salam:2008qg}, are available which perform PDF
evolution via an analogous procedure. {\tt APFEL} in particular, is
able to provide the so-called \textit{evolution operator}. The
evolution operator, is computed assuming that the DGLAP evolution can
be encoded in a single operator whose action on the intial scale PDFs
proceduces the final scale PDFs. Taking $Q_0$ as an initial scale and
$Q_\tau$ as a final scale, on the $x$-space grid one can write:
\begin{equation}\label{eq:fastPDFfinal_recalled}
  f_i(x_{\alpha},Q^2_\tau) = \sum_{k}
  \sum_\beta A^\tau_{\alpha\beta, ik}\;
  f_k(x_\beta,Q^2_0)\,, 
\end{equation}
where the operator $A$ is the solution of the DGLAP evolution equation
with baundary condition:
\begin{equation}
A^\tau_{\alpha\beta, ik} =
\delta_{\alpha\beta}\delta_{ik}\qquad\mbox{for}\qquad Q_\tau = Q_0\,.
\end{equation}

Given the evolution operator, we may replace the (general-scale) PDFs
used in the subprocess parton density Eq.~(\ref{eq:APPLsubproc}) with
their equivalent expressions c.f Eq.~(\ref{eq:fastPDFfinal_recalled})
like so:
\begin{equation}\label{eq:FK1}
\begin{array}{rcl}
F^{(s)}_{\alpha\beta,\tau} &=&  \displaystyle \sum_{i,j} \sum_{k,l}
                               \sum_{\delta,\gamma} C^{(s)}_{ij}
                               \left[  A^\tau_{\alpha\delta ik}\;
                               f_k(x_\delta,Q^2_0) A^\tau_{\beta\gamma
                               jl}\; f_l(x_\gamma,Q^2_0) \right]\;\;\;
  \\
\\
&=& \displaystyle \sum_{k,l}\sum_{\delta,\gamma}
\widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta}
f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0)\,,
\end{array}
\end{equation}
where the object:
\begin{equation}
  \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta} =
  \sum_{i,j} C^{(s)}_{ij} A^\tau_{\alpha\delta ik}
  A^\tau_{\beta\gamma jl}
\end{equation}
combines the operations of subprocess density construction and PDF
evolution. Going further it is possible to then use the expression for
subprocess parton densities in Eq.~(\ref{eq:FK1}) in the full
cross-section calculation:
\begin{equation}
\sigma_{pp\to X} = \sum_{k,l}\sum_{\delta,\gamma}\sum_p
\sum_{s} \sum_{\alpha,\beta}
\sum_{\tau} 
\alpha_s^{p+p_{\rm LO}}(Q^2_\tau)W_{\alpha\beta,\tau}^{(p)(s)} \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta}
f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0)\,.
\end{equation}
Performing some further contractions it is possible to obtain an
extremely compact expression for the calculation of the cross section
in question, in terms of only the initial-scale PDFs and summing only
over the initial scale interpolating $x$-grid and the incoming parton
flavors:
\begin{equation}\label{eq:FK}
  \sigma _{pp\to X} = \sum_{k,l}\sum_{\delta,\gamma} 
  \widetilde{W}_{kl,\delta\gamma} \,f_k(x_\delta,Q^2_0) f_l(x_\gamma,Q^2_0),
\end{equation}
where the object:
\begin{equation}\label{eq:FKTable}
  \widetilde{W}_{kl,\delta\gamma} = \sum_p^{N_\alpha} \sum_{s}^{N_{\mathrm{sub}}} \sum_{\alpha,\beta} \sum_{\tau}
\alpha_s^{p+p_{\rm LO}}(Q^2_\tau)  W_{\alpha\beta,\tau}^{(p)(s)} \widetilde{C}^{(s),\tau}_{kl,\alpha\beta\gamma\delta}
\end{equation}
is referred to here as an {\tt FK} table, and combines the information
stored in {\tt APPLgrid}-style interpolated weight grids with
analogously interpolated DGLAP evolution operators. This combination
enables for a maximally computationally-efficient expression for the
calculation of observables at hadron colliders under PDF variation.

\subsection{Features and limitations of {\tt FK} tables}
The {\tt FK} product of Eqn.~\ref{eq:FK} differs with respect to the product in Eqn.~\ref{eq:applconv} in several notable ways. Firstly 
the typical {\tt APPLgrid} or {\tt FastNLO} products require PDFs at a general scale $Q^2$, which implies performing PDF evolution e.g Eqn.~\ref{eq:fastPDFfinal_recalled} for every variation of the PDFs at the input scale during the fit. However in the {\tt FK} convolution this evolution
is pre-cached at the stage of {\tt FK} table generation, requiring only initial-scale PDFs as an input at fitting time. This pre-caching of the evolution
also removes the need for performing the sums over hard scale and perturbative order during the fit, further reducing the number of operations required.
The {\tt FK} convolution also benefits from the typically reduced number of active partonic modes at the fitting scale, with the sum over flavours in Eqn.~\ref{eq:FK} being limited to those directly used in the fit. Having reduced the calculation to such a simple form, it is also possible to straightforwardly
apply standard and widely-available computational tools such as multi-threading through e.g OpenMP or Single Instruction Multiple Data (SIMD) operations e.g SSE or AVX to further reduce computational cost.

While these features provide significant performance enhancements, the {\tt FK } table format is not suitable as a complete replacement for tools
such as {\tt APPLgrid}. The pre-computation of the PDF evolution necessarily means that all theory parameters such as perturbative order, strong coupling and factorization/renormalization scales are inextricably embedded in each {\tt FK} table. In order to perform PDF fits including variations of these parameters, multiple {\tt FK } tables must be computed, each with different theory settings. While performing such a re-calculation directly from Monte Carlo codes
would be exceptionally time consuming, the data representation in {\tt APPLgrid} files allows for an efficient combination with varying theory parameters.

%\section{The {\tt APFELgrid} package}
% **** Not sure if this sort of section would be neccesary ***
%The {\tt APFELgrid} plugin generates {\tt FK} tables of the form
%expressed in Eq.~(\ref{eq:FKTable}) from weight grids supplied by {\tt
%  APPLgrid} files. It is supplied as a set of {\tt C++} routines designed to
%  supplement the PDF evolution library {\tt APFEL} with {\tt FK} table generation
%  capabilities. The plugin has a very light interface for performing the {\tt FK} combination,
%  and relies on the native {\tt APFEL} API for setting the parameters of the PDF evolution,
%  including the structure of the initial scale $x$ grid.
%*** Would now discuss e.g SIMD implementation etc in advance of looking at the results in next section ***
\section{Performance benchmarks}
\label{sec:benchmark}

In this section we shall examine the performance
differences between the two expressions for cross-section prediction
Eq.~(\ref{eq:applconv}) ({\tt APPLgrid}) and Eq.~(\ref{eq:FK}) ({\tt FK}).
%
To provide a comprehensive benchmark, we consider here a wide range of
processes including LHC and Tevatron electroweak vector boson
production
measurements~\cite{Aaij:2012mda,Aaij:2012vn,Chatrchyan:2013mza,Chatrchyan:2013uja,Chatrchyan:2012xt,Aad:2013iua,Aad:2011fp,Aad:2011dm,Aaltonen:2010zza},
$t\bar{t}$ total
cross-sections~\cite{ATLAS:2012aa,ATLAS:2011xha,TheATLAScollaboration:2013dja,Chatrchyan:2013faa,Chatrchyan:2012bra,Chatrchyan:2012ria},
double-differential Drell-Yan
cross-sections~\cite{Chatrchyan:2013tia,CMS:2014jea} and inclusive jet
data~\cite{Chatrchyan:2012bja,Aad:2011fc,Aad:2013lpa,Abazov:2007jy}. Predictions
are performed over a wide range of kinematics, for a total of 52 source
{\tt APPLgrid} files corresponding to the majority of available LHC and Tevatron datasets of applicability to PDF determination.
While the source {\tt APPLgrid} files have varying grid densities in $x$ and $Q^2$, for the purposes of comparison the corresponding {\tt FK} tables are produced
consistently with 30 points in $x$, and at an initial scale below the charm threshold, therefore with seven active partonic species.

In Fig.~\ref{fig:timings} we compare the average time taken per datapoint
for the {\tt FK} and {\tt APPLgrid} calculations, for all of the 52
tables. We show timings for the {\tt FK} calculation in four
different configurations: AVX-OpenMP 2x (2 CPU cores), AVX, SSE3 and
standard double precision dot product. Due to the inherent structural differences between the {\tt FK} and {\tt APPLgrid} procedures,
results from the {\tt FK} calculation are systematically faster than those from {\tt APPLgrids}.
In particular, when comparing {\tt FK} AVX to {\tt APPLgrid} timings we obtain minimally a factor
of ten improvement in speed for electroweak vector boson production and
a maximum factor of 2000 improvement in predictions for inclusive jet data. Across all processes
and kinematic regions we observe significant performance improvements from using the {\tt FK} calculation
even in the maximally compatible, non-SIMD case.

While sheer computational speed is typically the primary consideration when computing observables
in a PDF fit, other factors such as table size in the filesystem and memory, along with the computational cost of
pre-computing {\tt FK} tables must be considered. Indeed, the computation of the {\tt FK} table in Eqn.~\ref{eq:FKTable}
requires a great deal of operations which can be time consuming, particuarly in the case of source {\tt APPLgrids} with very high
interpolation precision.

In Fig.~\ref{fig:performance} we examine the {\tt FK} table generation time with {\tt APFELgrid}, {\tt FK} table file size and memory usage of the {\tt FK} tables arising from the same source {\tt APPLgrids} as discussed in Fig.~\ref{fig:timings}. 
When examining the table generation time per point, we observe timings from a few milliseconds to
3.5 minutes per point, with differences arising from the varying grid densities used in the input {\tt APPLgrid} files. 
In terms of the grid size on disc, {\tt FK} tables are typically larger than their corresponding {\tt APPLgrids}, primarily as the {\tt FK} file format is encoding in plain text for compatibility whereas {\tt APPLgrids} are expressed in binary as {\tt ROOT} files. However when
measuring the in-memory resident set size used by the both convolution procedures, the amount of memory required by {\tt FK} tables is systematically less than {\tt APPLgrid} for all processes considered here by at least 75\%.

%In conclusion, the great performance benefits obtained with {\tt
%APFELgrid} endorse the importance of this tool not only in PDF fits
%but also in any application requiring a large number of convolutions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[scale=0.6]{plots/t0a}
\caption{\small Performance comparisons between {\tt APFELgrid} with
  AVX-OpenMP 2x, AVX, SSE3, double precision convolution and {\tt
    APPLgrid} convolution time per point and process.}
\label{fig:timings}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering                                      
  \includegraphics[scale=0.6]{plots/t0b}
\caption{\small {\tt APFELgrid} generation time per point and
  comparison to {\tt APPLgrid} for grid size and resident set size
  (RSS).}
\label{fig:performance}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusion}\label{sec:conclusion}

In this work we have presented {\tt APFELgrid}, an efficient
convolution interface to hadronic observables for fits of parton
densities. We have demonstrated that employing the so-called {\tt
  FastKernel} method, it is possible to convert an {\tt APPLgrid}
weight table into a derived format, referred to as {\tt FKtable},
which includes PDF and $\alpha_s$ evolution as provided by {\tt
  APFEL}.  Such a procedure allows one to obtain a very
computationally-efficient expression for the calculation of hadronic
cross sections, in terms of only the initial-scale PDFs and summing
only over the initial scale interpolating grid and the incoming parton
flavors. In addition, the simple structure of the resulting
convolution makes {\tt FKtables} particularly suitable to make
efficient use of acceleration and parallelisation techniques like {\tt
  SIMD} and {\tt OpenMP}.

We have shown that the numerical convolution of an {\tt APFELgrid}
tables is typically much faster than the corresponding {\tt
  APPLgrid}. In particular, the {\tt FastKernel} method provides a
great performance improvement with respect to {\tt APPLgrid} which is
further enhanced by the employment of any of the acceleration
techniques available in {\tt APFELgrid}. Finally, while the size of an
{\tt FKtable} is often larger than the corresponding {\tt APFELgrid}
file, we have shown that the resident set size memory required by {\tt
  APFELgrid} is typically smaller than {\tt APPLgrid} by at least
75\%.

The substantial speed improvement with respect to {\tt APPLgrid} in
association with the reduction of the resident set size memory, makes
the {\tt APFELgrid} code a valuable tool for modern PDF fits which
usually incluse thousands of hadronic data points and that thus
heavily rely on fast interfaces. We stress that {\tt APFELgrid} is not
a replacement for {\tt APPLgrid}, on which it depends, but it is
rather a specialised tool conceived to ease the inclusion of hadronic
data into PDF fits. In fact, the price of a better performance is the
fact that, by themselves, {\tt FKtables} cannot perform any parameter
variation other than that of PDFs. For example, variations of
$\alpha_s$ can be performed only by re-computing the {\tt FKtables}.

The {\tt APFELgrid} code is publicly available from the webpage:
\begin{center}
\url{https://github.com/nhartland/APFELgrid}
\end{center}
where intructions to install and run codes are also provided.

\section*{Acknowledgements}

We thanks the NNPDF Collaboration members for the support and
motivation they gave us to write this work. V.~B. and N.~H. are
supported by an European Research Council Starting Grant ``PDF4BSM''.
S.~C. is supported by the HICCUP ERC Consolidator grant (614577).

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix

%% \section{}
%% \label{}

%% References
%%
%% Following citation commands can be used in the body text:
%% Usage of \cite is as follows:
%%   \cite{key}         ==>>  [#]
%%   \cite[chap. 2]{key} ==>> [#, chap. 2]
%%

%% References with bibTeX database:
\section*{References}
\bibliographystyle{elsarticle-num}
\bibliography{paper}

%% Authors are advised to submit their bibtex database files. They are
%% requested to list a bibtex style file in the manuscript if they do
%% not want to use elsarticle-num.bst.

%% References without bibTeX database:

% \begin{thebibliography}{00}

%% \bibitem must have the following form:
%%   \bibitem{key}...
%%

% \bibitem{}

% \end{thebibliography}

\end{document}

%%
%% End of file